A Recipe for Scavenging - the natural history of a behaviour

### 2016-03-29 natural history of scavenging.tex
Minor edits to locomotion part

Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative to some extent \citep{beasley2015vertebrates}.
Therefore, we can use hyenas as examples of efficient terrestrial scavengers to compare with other forms. 
In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
The order Carnivora saw its origins in the Middle Eocene (56-33.9 Million years ago - Mya) where it split into the Caniforma and Feliforma. %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ... when it split into Caniforma and Feliforma." But I think we can leave it out because, god, it sounds pedantic!
So we can trace efficient terrestrial movement by carnivores from this point on.
Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene. %TG: ewgh... Thomas does not entirely agree. But then again, until now, reviewers didn't agreed with Thomas.

Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
Dinosaurs in general, were the dominant terrestrial forms for most of the Mesozoic Era. 
where the theropods ranged from the chicken-sized to the whale-sized, all of which were bipedal.
They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
Of relevance, are the questions that still persist about their metabolism with the latest evidence suggesting they were mesothermic i.e. intermediate to ecto- and endotherms \citep{grady2014evidence}. 
We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}. %TG: and cite "Kane et al in press" no? I think these people did some kind of work on that matter.
Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.

Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs. %TG: I will down that "dominance" term. Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
It is in the Permian, almost 300 millions years ago, that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have had the ability to range widely, a vital component in seeking out carrion.

>>>>

Today, terrestrial scavenging in the mammals is probably best known in an African context where hyenas, jackals and lions all take sizable proportions of carrion in their diet.
In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
Yet, no contemporary terrestrial vertebrate exists as an obligate scavenger but most if not all are facultative to some extent \citep{beasley2015vertebrates}.
Therefore, we can use hyenas as examples of efficient terrestrial scavengers to compare with other forms. 
In terms of locomotion, they employ a characteristic ``rocking horse gait"  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
The order Carnivora saw its origins in the Middle Eocene (56-33.9 Million years ago - Mya) where it split into the Caniforma and Feliforma. %TG: a phylo side note: the correct sentence would be the "The stem Carnivora saw its origins ... when it split into Caniforma and Feliforma." But I think we can leave it out because, god, it sounds pedantic!
So we can trace efficient terrestrial movement by carnivores from this point on.
Certainly, it is profitable to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene.

Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
They were the dominant terrestrial forms for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
Of relevance, are the questions that still persist about their metabolism with the latest evidence suggesting they were mesothermic i.e. intermediate to ecto- and endotherms \citep{grady2014evidence}. 
We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}. %TG: and cite "Kane et al in press" no? I think these people did some kind of work on that matter.
Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.

Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs. %TG: I will down that "dominance" term. Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
It is during the Permian (298.9 - 252.17 Mya) %, almost 300 millions years ago, %TG: Being pedantic here but Permian is just *not* 300 Mya... As a side note for the dates, I say we just use the geological epochs/periods/whatever since they are the actual precise dating of the fossil occurence (and they sound better). I suggest we just add the dates in brackets to give an indication.
that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have had the ability to range widely, a vital component in seeking out carrion.
